\chapter{Linear Algebra}

\section{Norm}
\[
    \|\vx\|_p = \biggl(\sum_i |x_i|^p\biggr)^{\frac{1}{p}}, p \in \R\ and\ p \geqslant 1
\]

范数是满足以下性质的任意函数：
\begin{itemize}
    \item $f(\vx) = 0 \Rightarrow \vx = 0$
    \item $f(\vx + \vy) \leqslant f(\vx) + f(\vy)$
    \item $\forall \alpha \in \R, f(\alpha \vx) = |\alpha|f(\vx)$
\end{itemize}

\section{Eigenvalues and Eigenvectors}
\begin{equation}
    Au = \lambda u
\end{equation}
特征值$\lambda$代表线性变化的伸缩倍数，特征向量$u$代表变换的方向

\section{Singular value decomposition}
\textbf{定义} 对于$A \in C^{m \times n}$，$rank(A) = r$，矩阵$A^HA$的特征值为
$\lambda_1 \geqslant \lambda_2 \geqslant ... \geqslant \lambda_r > 0$，
$\lambda_{r+1} = \lambda_{r+2} = ... = \lambda_{n} = 0$，称正数$\sigma_i = \sqrt{\lambda_i}(i = 1,2,...,r) $
为矩阵$A$的\textbf{奇异值}
\\
\begin{equation}
    \begin{split}
        A
        &= U \Sigma V^H \\
        &=
        \begin{pmatrix}
            u_1 & u_2 & \cdots & u_m
        \end{pmatrix}
        \left(
            \begin{array}{cccc:c}
                \sigma_1&           &           &           & \\
                        & \sigma_2  &           &           & 0\\
                        &           & \ddots    &           & \\
                        &           &           & \sigma_r  & \\
                        \hdashline
                        &           & 0         &           & 0
            \end{array}
        \right)
        \begin{pmatrix}
            v_1^H \\
            v_2^H \\
            \cdots \\
            v_n^H
        \end{pmatrix} \\
        &= \sigma_1 \boldsymbol{u}_1 \boldsymbol{v}_1^H + \sigma_2 \boldsymbol{u}_2 \boldsymbol{v}_1^H + \cdot\cdot\cdot + \sigma_r \boldsymbol{u}_r \boldsymbol{v}_r^H
    \end{split}
\end{equation}

\subsection{Dimensionality Reduction}
将图片作为矩阵进行奇异值分解，提取前n个奇异值，则可以达到图像压缩的目的.


\subsection{Matrix}
\href{https://math.stackexchange.com/questions/42649/why-are-invertible-matrices-called-non-singular}{Why are invertible matrices called 'non-singular'?}

If you take an $n \times n$ matrix "at random", then it will almost certainly be invertible. 
That is, the generic case is that of an invertible matrix, the special case is that of a matrix that is not invertible.

For example, a $1 \times 1$ matrix(with real coefficients) is invertible if and only if it is not the $\mathbf{0}$ matrix; 
for $2 \times 2$ matrices, it is invertible if and only if the two rows do not lie in the same line through the origin; 
for $3 \times 3$, if and only if the three rows do not lie in the same plane through the origin; etc.

So here, "singular" is not being taken in the sense of "single", but rather in the sense of "special", "not common". 
See the dictionary definition: it includes "odd", "exceptional", "unusual", "peculiar".

The noninvertible case is the "special", "uncommon" case for matrices. 
It is also "singular" in the sense of being the "troublesome" case (you probably know by now that when you are working with matrices, the invertible case is usually the easy one).

\textbf{满秩矩阵一定是可逆矩阵}